### 2020-08-30 paper/dcpo_gender_roles(Intro).Rmd
I add Intro portion to the previous rmd.

---

output:

  pdf_document:

    citation_package: natbib

    keep_tex: true

    fig_caption: true

    latex_engine: pdflatex

    template: svm-latex-ms2.tex

title: "[dcpo_gender_roles]: A Dataset of Dynamic Comparative Public Opinion Toward Egalitarian Gender Roles"

thanks: "Corresponding author: [frederick-solt@uiowa.edu](mailto:frederick-solt@uiowa.edu).  Current version: `r format(Sys.time(), '%B %d, %Y')`."

author:

abstract: ""

keywords: ""

date: "`r format(Sys.time(), '%B %d, %Y')`"

fontsize: 11pt

spacing: double

bibliography: \dummy{`r file.path(getwd(), list.files(getwd(), ".bib$", recursive = TRUE))`}

# csl: https://raw.githubusercontent.com/citation-style-language/styles/master/american-political-science-association.csl

biblio-style: apsr

citecolor: black

linkcolor: black

endnote: no

header-includes:

      - \usepackage{array}

      - \usepackage{caption}

      - \usepackage{graphicx}

      - \usepackage{siunitx}

      - \usepackage{colortbl}

      - \usepackage{multirow}

      - \usepackage{hhline}

      - \usepackage{calc}

      - \usepackage{tabularx}

      - \usepackage{threeparttable}

      - \usepackage{wrapfig}

nocite: |

  @Solt2020a

---

```{r setup, include=FALSE}

options(tinytex.verbose = TRUE)

knitr::opts_chunk$set(cache = TRUE, echo = FALSE, fig.width=6.5, fig.height=2.5, cache.extra = tools::md5sum(here::here("data", "dcpo_input_raw.csv")))

# If `DCPOtools` is not yet installed:

# remotes::install_github("fsolt/DCPOtools")

library(DCPOtools)

library(DCPO)

library(tidyverse)

set.seed(324)

```

#Introduction
Much of the comparative public opinion literature that does consider gender places its importance at the individual level. Such scholarship focuses on how men and women express significantly different opinions in various issue areas such as public policy preferences [@kellstedt2010macro], the use of military force [@brooks2011war; @eichenberg2016gender], and female political leadership [@badas2019michelle] to name a few. While such research has produced interesting and important findings by asking how gender affects public opinion, it ignores a second equally relevant piece of the puzzle that our data is better equipped to address – that is, how does public opinion affect gender? This analytical reframing is important, because without it, the literature loses sight of gender as not only a variable indicating part of an individual’s identity, but also as a macrolevel process that groups of people produce, experience, and challenge as part of their political lives. By measuring cross-national time-variant attitudes toward gender egalitarianism, our data positions researchers to ask questions about how public opinion relates to the macrolevel processes of gender.
Gender egalitarian attitudes have been grown over decades not only in developed countries but also in developing societies [@boehnke2011gender]. With the affluent variance in the gender egalitarian attitudes around the world, scholars propose innovative research questions about the influence of gender egalitarian attitudes and its determinants. For example, the association between gender egalitarian attitudes and women’s political behavior [@krauss1974political], gender inequalities [@bianchi2012housework], women’s descriptive representation in parliaments [@paxton2003women] have been examined. However, the lack of cross-national time-series index prevents us from unveiling the cause and effect of gender egalitarian attitudes. Very few cross-national time-series measures of gender egalitarianism exist and there is no consensus on the measures. Even though a gender egalitarian index developed by @mcdaniel2008measuring  exists, it only deals with limited number of countries (n=38) and with the fourth wave of World Value Survey (WVS) conducted between 1999 and 2004.
The newly developed gender egalitarian index enables scholars to evaluate the generalizability of existing theories with large-N analysis, test the cause and effect of gender egalitarian attitudes, build new theories with the preliminary statistical analysis, etc. As the very first step, the index provides an opportunity to descriptively examine how the attitudes have been changed over decades and why there are variations among countries with different political institutions, culture, religion, and other country specific factors. Also, we can test the long-lasting question suggested by @krauss1974political about whether the differences in political behavior such as protests, voting, political volunteerism, and feminist movement of men and women are attributed to biological differences or public’s views on the gender role. Moreover, the index makes it possible to unveil the relationship between gender egalitarianism and women’s mobilization and representation [@erickson2002gender; @erickson2002gender; @klein2013gender; @paxton1997women] with our index. Further, the broad time span of our index enables scholars to test causality and the direction of the relationship.
	One example of how our data might be used to investigate these macrolevel processes is researching how public opinion regarding gender roles affects the makeup of national militaries. The comparative public opinion literature has completely neglected this connection, despite scholarly interest in how public opinion affects policy decisions [@stimson1995dynamic; @lax2009gay; @Wwlezien2017public], the relationship between public opinion and political institutions [@bartels2009explaining; @grieco2011let; @nir2015institutions], and how gender affects public opinion toward military actions [@brooks2011war; @eichenberg2016gender; @lizotte2017gender]. If there exists a strong correlation between countries that largely hold traditionalist views toward gender roles and countries that exclude women from participating in the state military, then this would support a theory that national military policies tend to be congruent with public opinion. One could theorize that this is because of incentive structures military policy-makers face, and perhaps the level of congruence between public opinion and military policy is conditioned by other factors such as regime type. 
However, it would be just as interesting to discover that there is no correlation between public opinion on gender roles and women’s inclusion/exclusion from state militaries. If the countries with largely egalitarian views toward gender roles are no more likely to incorporate women into the national military than more traditionalist countries, then this would reveal important patterns in the gendered makeup of national militaries. Perhaps these typically male-dominated institutions have shifted over time from being male-dominated as a norm to being male-dominated as an active resistance to public demands for change – in which case, national militaries have developed an identity as not just masculine institutions, but as preservers of traditional masculinity. Regardless of what results might unfold from this avenue of research, using our data on attitudes toward gender egalitarianism to test theories regarding macroprocesses of gender will help move the comparative public opinion literature beyond just identifying how male and female survey respondents differ from one another.

# The Source Data on Gender Egalitarian Attitudes

```{r dcpo_input_raw, include=FALSE, cache.extra = tools::md5sum(here::here("data-raw", "surveys_egr.csv"))}

surveys_egr <- read_csv(here::here("data-raw", "surveys_egr.csv"),

                        col_types = "cccc")

dcpo_input_raw <- DCPOtools::dcpo_setup(vars = surveys_egr,

                                        datapath = here::here("..", "data", "dcpo_surveys"),

                                        file = here::here("data", "dcpo_input_raw.csv"))

```

```{r egr_summary_stats}

dcpo_input_raw1 <- dcpo_input_raw %>%

  with_min_yrs(2) %>%

  with_min_cy(5) %>%

  group_by(country) %>%

  mutate(cc_rank = n()) %>%

  ungroup() %>%

  arrange(-cc_rank)

n_surveys <- surveys_egr %>%

  distinct(survey) %>%

  nrow()

n_items <- dcpo_input_raw1 %>%

  distinct(item) %>%

  nrow()

n_countries <- dcpo_input_raw1 %>%

  distinct(country) %>%

  nrow()

n_years <- as.integer(summary(dcpo_input_raw1$year)[6]-summary(dcpo_input_raw1$year)[1])

total_cy <- {n_countries * n_years} %>%

  scales::comma()

year_range <- paste("from",

                    summary(dcpo_input_raw1$year)[1],

                    "to",

                    summary(dcpo_input_raw1$year)[6])

n_cyi <- dcpo_input_raw1 %>%

  distinct(country, year, item) %>%

  nrow() %>%

  scales::comma()

```

The first step towards remedying this problem is collecting the available public opinion data on gender egalitarian attitudes.

We draw on data from `r n_surveys` survey datasets, in which we identified `r n_items` distinct relevant survey items that were asked in no fewer than five country-years.

Together, these survey items were asked in `r n_countries` different countries in at least two time points over `r n_years` years, `r year_range`, yielding a total of `r n_cyi` country-year-item observations.

Considering that observations for every year in each country surveyed would number `r total_cy` and so a complete set of country-year-items would encompass `r {n_countries * n_years * n_items} %>% scales::comma()` observations, the available data is clearly very, very sparse.

```{r obs_by_item, fig.height=3, fig.cap = "Items with the Most Observations in the Source Data \\label{items_by_obs}"}

dcpo_input_raw1 %>%

  distinct(country, year, item) %>%

  count(item) %>%

  arrange(desc(n)) %>%

  head(12) %>%

  ggplot(aes(forcats::fct_reorder(item, n, .desc = TRUE), n)) +

  geom_bar(stat = "identity") +

  theme_bw() +

  theme(axis.title.x = element_blank(),

        axis.text.x  = element_text(angle = 90, vjust = .45, hjust = .95)) +

  ylab("Country-Years Observed")

polileader4_cy <- dcpo_input_raw1 %>% filter(item == "polileader4") %>% distinct(country, year) %>% nrow()

polileader4_surveys <- dcpo_input_raw1 %>% filter(item == "polileader4") %>% distinct(survey) %>% pull(survey)

housewifefull4_cy <- dcpo_input_raw1 %>% filter(item == "housewifefull4") %>% distinct(country, year) %>% nrow()

```

Figure&nbsp;\ref{items_by_obs} displays in how many country-years each of the twelve most-commonly asked items are available.

The `polileader4` item, which asks respondents whether they strongly agree, agree, disagree, or strongly disagree with the statement "On the whole, men make better political leaders than women do," was the most frequently asked question in the data we collected.

Employed by the Americas Barometer, the Arab Barometer, the Eurobarometer, the Latinobarómetro, the Pew Research Center, and the World Values Survey, this question was asked in a total of `r polileader4_cy` different country-years.

That this constitutes only `r round(polileader4_cy*100/(n_countries * n_years))`% of the `r total_cy` total possible country-years covered---and remember, `polileader4` is the _most common_ survey item---again underscores just how sparse the available public opinion data is on this topic.

Which countries are the most data-rich?

Figure&nbsp;\ref{countries_by_obs} below shows the dozen countries with the highest count of country-year-item observations.

```{r obs_by_country, fig.cap = "Countries with the Most Observations in the Source Data \\label{countries_by_obs}"}

dcpo_input_raw1 %>%

  mutate(country = if_else(stringr::str_detect(country, "United"),

                           stringr::str_replace(country, "((.).*) ((.).*)", "\\2.\\4."),

                           country)) %>%

  distinct(country, year, item) %>%

  count(country) %>%

  arrange(desc(n)) %>%

  head(12) %>%

  ggplot(aes(forcats::fct_reorder(country, n, .desc = TRUE), n)) +

  geom_bar(stat = "identity") +

  theme_bw() +

  theme(axis.title.x = element_blank(),

        axis.text.x  = element_text(angle = 90, vjust = .45, hjust = .95)) +

  ylab("Year-Items Observed")

# uk_nas <- {1 + max(swiid_source$year) - min(swiid_source$year) -

# {swiid_source %>% filter(country == "United Kingdom") %>% pull(year) %>% unique() %>% length()}} %>%

# wordify_numeral()

#

# iran_obs <- swiid_source %>% filter(country == "Iran") %>% pull(year) %>% unique() %>% length()

#

# med_cy <- swiid_source %>% count(country, year) %>% count(country, name = "nn") %>% pull(nn) %>% median() %>% wordify_numeral()

```

```{r years_by_country, fig.cap="Country-Year Coverage in the Source Data \\label{countries_by_years}"}

library(patchwork)

cby_plot <- dcpo_input_raw1 %>%

  mutate(country = if_else(stringr::str_detect(country, "United"),

                           stringr::str_replace(country, "((.).*) ((.).*)", "\\2.\\4."),

                           country),

         country = stringr::str_replace(country, "South", "S.")) %>%

  distinct(country, year) %>%

  count(country) %>%

  arrange(desc(n)) %>%

  head(12) %>%

  ggplot(aes(forcats::fct_reorder(country, n, .desc = TRUE), n)) +

  geom_bar(stat = "identity") +

  theme_bw() +

  theme(axis.title.x = element_blank(),

        axis.text.x  = element_text(angle = 90, vjust = .45, hjust = .95)) +

  ylab("Years Observed")

ybc_plot <- dcpo_input_raw1 %>%

  distinct(country, year) %>%

  count(year, name = "nn") %>%

  ggplot(aes(year, nn)) +

  geom_bar(stat = "identity") +

  theme_bw() +

  theme(axis.title.x = element_blank(),

        axis.text.x  = element_text(angle = 90, vjust = .45, hjust = .95)) +

  ylab("Countries Observed")

cby_plot + ybc_plot

```

```{r dcpo_input}

dcpo_input <- DCPOtools::format_dcpo(dcpo_input_raw1,

                                     scale_q = "polileader4",

                                     scale_cp = 2)

save(dcpo_input, file = here::here("data", "dcpo_input.rda"))

```

# Estimating Gender Egalitarianism

The DCPO model is estimated using the `DCPO` package for R [@Solt2020a], which is written in the Stan probabilistic programming language [@StanDevTeam2019a; @StanDevTeam2019b].

```{r dcpo, eval=FALSE}

iter <- 4000

dcpo_output <- dcpo(dcpo_input,

                    iter = iter,

                    chains = 4,

                    thin = iter/500, # this yields 250 draws per chain, 1000 draws total

                    pars = c("sd_delta","sd_theta_evolve", "sd_sigma_evolve", "sigma","phi","beta","alpha","delta","theta","y_r_pred","log_lik"))

save(dcpo_output, file = here::here("data", "dcpo_output.rda"))

dcpo_theta <- rstan::extract(dcpo_output, pars = "theta")

save(dcpo_theta, file = here::here("data", "dcpo_theta.rda"))

dcpo_sigma <- rstan::extract(dcpo_output, pars = "sigma")

save(dcpo_theta, file = here::here("data", "dcpo_sigma.rda"))

dcpo_y_r_pred <- rstan::extract(dcpo_output, pars = "y_r_pred")

save(dcpo_input, dcpo_y_r_pred, file = here::here("data", "dcpo_y_r_pred.rda"))

```

```{r cs_plot, fig.cap="Egalitarian Gender Attitudes, 2020 \\label{cs_2020}", fig.height=8, warning=FALSE}

load(here::here("data", "dcpo_input.rda"))

load(here::here("data", "dcpo_output.rda"))

theta_results <- DCPO::summarize_dcpo_results(dcpo_input,

                                              dcpo_output,

                                              "theta")

sigma_results <- DCPO::summarize_dcpo_results(dcpo_input,

                                              dcpo_output,

                                              "sigma")

p1_data <- theta_results %>%

    group_by(country) %>%

    top_n(1, year) %>%

    ungroup() %>%

    arrange(-mean) %>%

    mutate(half = ntile(mean, 2))

p1_data_a <- p1_data %>% filter(half==2) %>% mutate(ranked = as.factor(row_number(mean)))

p1_data_b <- p1_data %>% filter(half==1) %>% mutate(ranked = as.factor(row_number(mean)))

p1a <- ggplot(p1_data_a, aes(x = mean,

                             y = ranked)) +

    geom_segment(aes(x = `10%`, xend = `90%`,

                     y = ranked, yend = as.factor(ranked)),

                 na.rm = TRUE) +

    geom_point(fill = "black", shape = 21, size = 1.1, na.rm = TRUE) +

    theme_bw() + theme(legend.position="none",

                       axis.text.x  = element_text(size=8),

                       axis.text.y  = element_text(size=8),

                       axis.title.x = element_text(face="bold", size=8),

                       legend.text = element_text(size = 7),

                       legend.title = element_text(size=7, face = "bold"),

                       legend.key.size = unit(.5, "line"),

                       legend.background = element_rect(linetype = "solid",

                                                        color = "grey80",

                                                        size = .25),

                       legend.key = element_rect(colour = "white")) +

    scale_y_discrete(breaks = p1_data_a$ranked, labels=p1_data_a$country) +

    coord_cartesian(xlim=c(0, 1)) +

    labs(x = NULL, y = NULL)

p1b <- ggplot(p1_data_b, aes(x = mean,

                             y = ranked)) +

    geom_segment(aes(x = `10%`, xend = `90%`,

                     y = ranked, yend = ranked),

                 na.rm = TRUE) +

    geom_point(fill = "black", shape = 21, size = 1.1, na.rm = TRUE) +

    theme_bw() + theme(legend.position="none",

                       axis.text.x  = element_text(size=8),

                       axis.text.y  = element_text(size=8),

                       axis.title.x = element_text(face="bold", size=8)) +

    scale_y_discrete(breaks = p1_data_b$ranked, labels=p1_data_b$country) +

    coord_cartesian(xlim=c(0, 1)) +

    labs(x = NULL, y = NULL)

p1a + p1b

```

# Conclusion

\pagebreak