### 2022-03-21 paper/response_memo.Rmd
address #31

---
output: 
  pdf_document:
    citation_package: natbib
    keep_tex: true
    fig_caption: true
    latex_engine: xelatex
    template: ../svm-latex-ms2.tex
title: |
    | Public Gender Egalitarianism: A Dataset of Dynamic Comparative
  Public Opinion Toward Egalitarian Gender Roles in the Public Sphere
    | Memo to Reviewers
date: "`r format(Sys.time(), '%B %d, %Y')`"
fontsize: 11pt
spacing: single
bibliography: dcpo_gender_roles.bib
biblio-style: apsr
citecolor: black
linkcolor: black
endnote: no
header-includes:
      - \usepackage{array}
      - \usepackage{caption}
      - \usepackage{graphicx}
      - \usepackage{siunitx}
      - \usepackage{colortbl}
      - \usepackage{multirow}
      - \usepackage{hhline}
      - \usepackage{calc}
      - \usepackage{tabularx}
      - \usepackage{threeparttable}
      - \usepackage{wrapfig}
---

```{r setup, include=FALSE}
options(tinytex.verbose = TRUE)

knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  dpi = 300)

# If you haven't install `DCPOtools`
# remotes::install_github("fsolt/DCPOtools")

if (!require(pacman)) install.packages("pacman")
library(pacman)
library(patchwork)
library(rsdmx)

# load all the packages you will use below 
p_load(
  dataverse, # data scraping
  DCPOtools,
  boot, plm, # analysis
  flextable, kableExtra, modelsummary, # tabulation
  gridExtra,
  latex2exp, # visualization
  rstan, # Bayesian estimation
  broom, tidyverse, janitor, # data wrangling
  dotwhisker # visualization
)

# Functions preload
set.seed(313)

get_legend<-function(myggplot){
  tmp <- ggplot_gtable(ggplot_build(myggplot))
  leg <- which(sapply(tmp$grobs, function(x) x$name) == "guide-box")
  legend <- tmp$grobs[[leg]]
  return(legend)
}

```

Thank you for your helpful comments, as well as your general enthusiasm for this piece.
We set out a list of the specific points raised in the reviews and our responses to them below, roughly in the order they appear in the text:

1. **Introduction** R2 asked if we were "only using cross-national surveys or whether you also incorporate single-country surveys."
We clarified that we are drawing on both cross-national and single-country surveys at the top of page 2.
We also revised the sentence on peak country coverage (at the bottom of p3) that prompted R2's question in an effort to avoid any confusion.

1. **Examining the Source Data on Public Gender Egalitarianism** R1 noted three spots in the first few pages where our language was respectively unclear, mistaken, and repeated.
We corrected each of these infelicities with gratitude.

1. **Estimating Public Gender Egalitarianism** R2 suggested that our description of the model could be made more accessible given the broad audience of the _BJPS_ and the work's potential for use in the classroom.
We agree entirely.
We revised these paragraphs carefully for readability, shifting much of the technical discussion to Appendix C and instead elaborating on how the model's terms address the twin limitations of the source survey data that we raise in the paper's previous section, their incomparability and their sparsity.

1. **Estimating Public Gender Egalitarianism** R1 requested "more description about how the estimates were obtained," specifically details about iterations, chains, warmup, thinning, and convergence.  We added a short paragraph with this information at p5.

Thank you once more for the opportunity to make revisions.
We think the paper is much stronger as a result of your comments, and we hope you agree.

# References

<div id="refs"></div>